---
title: "Project Bar Graph and Summary Table"
author: "Gunjan Joshi"
date: "10/2/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/gunja/Desktop/Hw/Data 106/Group_Project")

library(readr)
library(dplyr)
library(reshape2)
library(tidyr)
library(ggplot2)


medals <- read_csv("No_of_medals_fixed.csv")

participants <- read_csv("total_participants.csv")
#write.csv(Australia, "Australia.csv")
```

```{r }

asian_c = c("China", "India", "Japan", "Korea, South", "Vietnam")
north_a = c("United States", "Canada", "Jamaica", "Mexico", "Cuba")
south_a = c("Brazil", "Argentina", "Chile", "Colombia", "Venezuela")
aus_nz = c("Australia", "New Zealand")
eu = c("Germany", "Poland", "Sweden", "Bulgaria", "United Kingdom")
afr= c("South Africa", "Ghana", "Zimbabwe", "Kenya")

continents = medals %>%
  mutate(Continent = ifelse(`Country Name` %in% 
                              asian_c, "Asia", 
                                     ifelse(`Country Name` %in%
                                              north_a, "N.America",
                                            ifelse(`Country Name` %in% 
                                                     south_a, "S.America",
                                                   ifelse(`Country Name` %in% 
                                                            aus_nz, "Australia(&NZ)",
                                                          ifelse(`Country Name` %in% 
                                                                   eu, "Europe",
                                                                 "Africa"))))))
a = participants %>% 
  mutate(Continent = ifelse(`Country Name` %in% 
                              asian_c, "Asia", 
                                     ifelse(`Country Name` %in%
                                              north_a, "N.America",
                                            ifelse(`Country Name` %in% 
                                                     south_a, "S.America",
                                                   ifelse(`Country Name` %in% 
                                                            aus_nz, "Australia(&NZ)",
                                                          ifelse(`Country Name` %in% 
                                                                   eu, "Europe",
                                                                 "Africa"))))))
```

1. Create a relevant summary table. Make sure this table is not super long or wordy. Also
make sure the columns are renamed appropriately.

```{r message=FALSE}
final = continents %>% 
  mutate(number_medals = ifelse(is.na(number_medals), 0, number_medals)) 
z = final %>%group_by(Year, Continent) %>% 
  summarise(Total_Medals = sum(number_medals))
q = a %>%group_by(Year, Continent) %>% 
  summarise(Male_p = sum(Men), Female_p = sum(Women), All_Participant = sum(Total_participants))
ratio = q %>% mutate(`Women to Men participants` = Female_p / Male_p)


last = merge(z, ratio, by.x = c("Continent", "Year"), by.y = c("Continent", "Year"))

names(last)[4] <- "Men"
names(last)[5] <- "Women"
names(last)[3] <- "Total Medals"
names(last)[6] <- "Total Participants"
last

```



2. Create a relevant plot (either boxplot, barplot (live session) or scatterplot (pre-recorded)).
Make sure to add color, nice titles (maybe captions and subtitles) and edit the plot a little
(maybe change the colors, the transparency of the colors etc.)

```{r message=FALSE}


cont = final %>% group_by(Year, `Continent`) %>% summarise(Total_Medals = sum(number_medals))


cont %>% mutate(Year= factor(Year)) %>% 
  ggplot(aes(x =Year, y=Total_Medals, fill=Year, color= Year))+
  geom_bar(stat = "identity", position = "dodge", alpha= 0.85, color = "black")+
  facet_wrap(~Continent)+ 
  theme(axis.text.x=element_text(angle=45, hjust=1))+ 
  ggtitle("Total Medals Won in the Olympics")+
  xlab("Years (1992-2012)")+
  ylab("Number of Medals Won")+
  labs(subtitle = "Medals won by different Continents from 1992- 2012")
```
**DATASET FOR MODELS**

```{r message=FALSE}

gold = final %>% filter(Medal == "Gold")
gold2 = gold %>% group_by(Year, `Country Name`) %>% #, `GDP Per Capita`, HDI_Value, Population) 
                         summarise(`Total Gold Medals` = sum(number_medals))


p = final %>% group_by(`Country Name`, Year) %>% slice(1) %>% 
  select(Year, `Country Name`, `GDP Per Capita`, HDI_Value, Population)


r = a %>%group_by(Year, `Country Name`) %>% 
  summarise(Male_p = sum(Men), Female_p = sum(Women), All_Participant = sum(Total_participants))%>%
  mutate(`Women to Men participants` = Female_p / Male_p)

dataset = merge(p, r, by=c("Year", "Country Name"), all = TRUE)

dataset2 = merge(dataset, gold2, by=c("Year", "Country Name"),all = TRUE)

f = final %>%group_by(Year, `Country Name`) %>% 
  summarise(Total_Medals = sum(number_medals))
project = merge(dataset2, f, by=c("Year", "Country Name"),all = TRUE)

model_dataset = project %>% mutate(`Total Gold Medals` = ifelse(is.na(`Total Gold Medals`), 0, `Total Gold Medals`))

```


```{r}
Africa = last %>% filter(Continent == "Africa")
```


```{r}
N.America = last %>% filter(Continent == "N.America")
```

```{r}
Europe = last %>% filter(Continent == "Europe") 
S.America = last %>% filter(Continent == "S.America")
Asia = last %>% filter(Continent == "Asia")
Australia = last %>% filter(Continent == "Australia(&NZ)")
```

```{r}
model_dataset
```


```{r}
scatter_data = model_dataset %>% mutate(Continent = ifelse(`Country Name` %in% 
                              asian_c, "Asia", 
                                     ifelse(`Country Name` %in%
                                              north_a, "N.America",
                                            ifelse(`Country Name` %in% 
                                                     south_a, "S.America",
                                                   ifelse(`Country Name` %in% 
                                                            aus_nz, "Australia(&NZ)",
                                                          ifelse(`Country Name` %in% 
                                                                   eu, "Europe",
                                                                 "Africa"))))))
```


```{r}

bar = scatter_data %>% group_by(Year, `Continent`) %>% summarise(`Total Gold Medals` = sum(`Total Gold Medals`))




bar %>% mutate(Year= factor(Year)) %>% 
  ggplot(aes(x =Year, y= `Total Gold Medals`, fill= Continent))+
  geom_bar(stat = "identity", position = "dodge", alpha= 0.85,  color= "black")+
  ggtitle("Total Gold Medals Won in the Olympics")+
  xlab("Years (1992-2012)")+
  ylab(" Total Number of Gold Medals Won")+
  labs(subtitle = " Gold Medals won by different Continents from 1992- 2012")
```
```{r}
gdp_plot = scatter_data %>% group_by(Year, `Continent`) %>% summarise(`Average GDPPC` = mean(`GDP Per Capita`))
gdp_plot %>% mutate(Year= factor(Year)) %>% 
  ggplot(aes(x =Year, y= `Average GDPPC`, fill= Continent))+
  geom_bar(stat = "identity", position = "dodge", alpha= 0.85, color= "black")+
  ggtitle("Average GDPPC for different Continents")+
  xlab("Years (1992-2012)")+
  ylab(" Average GDPPC")+
  labs(subtitle = " Average GDP from 1992 - 2012")
```

